Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-02 introduction.tex
Major update to Section 1 and 2

The results are obtained by coupling the hyperbolic random graph to an infinite random graph $G_\Pcal(\alpha,\nu)$, which we define in Section~\ref{ssec:infinite_model}, and show that the limit of the local clustering coefficient and function for the hyperbolic random graph are given by those for the infinite model. To express these expressions we define for any $y \in \R_+$ 
\[
	\rho(y,k) = \Prob{\Po\left(\xi_{\alpha,\nu} e^{y/2}\right) = k},
\]
where $\xi_{\alpha,\nu} = (4\alpha \nu)/(2\alpha - 1)\pi$ and $\Po(\lambda)$ denotes a Poisson random variable with mean $\lambda$. In addition we define the (marginal) probability measures on $\Rcal := \R_+ \times \R$ by  
\[
	\eta_{y}(x^\prime, y^\prime) = \frac{\alpha \nu e^{-\frac{y}{2}}}{\xi_{\alpha,\nu} \pi} \, \ind{|x^\prime|<e^{(y+y^\prime)/2}} \, e^{-\alpha y^\prime}
\]	
and let
\begin{equation}\label{eq:def_delta_p}
	\Delta_{\Pcal}(y) = \iint_{\Rcal^2} T_\Pcal(y, x_1, x_2, y_1, y_2) \eta_{y}(x_1,y_1)\eta_{y}(x_2,y_2) \dd x_1 \dd x_2  \dd y_1  \dd y_2,
\end{equation}
where 
\[
	T_\Pcal(y, x_1, x_2, y_1, y_2) 
	= \ind{|x_1| \, \le \, e^{(y+y_1)/2}}\ind{|x_2| \, \le \, e^{(y+y_2)/2}}\ind{|x_1-x_2| \, \le \, e^{(y_1+y_2)/2}}.
\]

With these expressions we define the limiting local clustering coefficient as
\begin{equation}\label{eq:def_average_clustering_infinite_model}
	c_\infty := \alpha \int_0^\infty \Delta_{\Pcal}(y) (1 - \rho(y,0) - \rho(y,1)) e^{-\alpha y} \dd y,
\end{equation}
while the limiting local clustering function is defined as
\begin{equation}\label{eq:def_c_infty_k}
	c_\infty(k) = \frac{\int_0^\infty \rho(y,k) \Delta_\Pcal(y) e^{-\alpha y} \dd y}{\int_0^\infty \rho(y,k) e^{-\alpha y} \dd y}.
\end{equation}

Finally to express the local clustering coefficient and function in the infinite limit model we need a couple of special functions.

\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function 
%\[
%	B(a,b) = \int_0^1 t^{a-1}(1-t)^{b-1} \, dt
%\]
and $B^-(x,a,b)$ the lower incomplete beta-function
\[
	B^-(x,a,b) = \int_x^1 t^{a-1}(1-t)^{b-1} \, dt.
\]
In addition we write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function
\[
	\Gamma^+(q,z) = \int_z^\infty t^{q - 1} e^{-t} \dd t
\]
and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

@@

\PvdH{@All: I would be nice if we could show with simulations how close $c_{\H,n}$ actually is to $c_\infty$ for reasonably large $n$ in practice.}

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to $c_\infty$.

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0.
\]
\end{theorem}

Next we establish the full expression for the limiting clustering coefficient $c_\infty$ in terms of the special functions defined above.

\begin{theorem}[Exact expression for limiting clustering coefficient]\label{thm:exact_expression_c_infty}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then, if $\alpha \ne 1$
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
while for $\alpha = 1$,
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}. 
\end{theorem}

@@

We obtain similar results for the local clustering function. The first states that $c_{\H,n}(k)$ is asymptotically equivalent to $c_\infty(k)$ in an $L^1$ sense.

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:asymptotic_local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0.
\]
%where
%\[
%	C_{\alpha,\nu} = \begin{cases}
%		C_\alpha \left(\frac{2\alpha \nu}{\pi\left(\alpha - \frac{1}{2}\right)}\right)^{4\alpha - 2}
%		&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
%		\frac{2\alpha \nu}{\pi\left(\alpha - \frac{1}{4}\right)} &\mbox{if } \alpha = \frac{3}{4},\\
%		\frac{2\alpha \nu}{\pi\left(\alpha - \frac{3}{4}\right)} &\mbox{if } \alpha \ge \frac{3}{4}
%	\end{cases} 
%\]
%with $C_\alpha$ as defined in \eqref{eq:def_C_alpha}.
\end{theorem}

Next we give an exact expression for $c_\infty(k)$ in terms of the special functions defined earlier.

\begin{theorem}[Exact expression for limiting local clustering function]
\label{thm:exact_expression_local_clustering}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then, for $\alpha \ne 1$
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}

%\begin{align*}
%	c_\infty(k) 
%	&= \frac{\xi_{\alpha,\nu}^{4\alpha - 2}\Gamma^+(k-6\alpha + 2,\xi_{\alpha,\nu})}{\Gamma^+(k-2\alpha,\xi_{\alpha,\nu})}
%		\left(\frac{2^{-4\alpha - 1}(3\alpha - 1)}{\alpha(\alpha - 1)^2} + \frac{(\alpha - \frac{1}{2})B^-\left(\frac{1}{2},2\alpha + 1, 2\alpha - 2\right)}{2\alpha(\alpha - 1)}\right)\\
%	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{4\alpha - 2}\Gamma(2\alpha + 1)}{\Gamma^+(k-2\alpha,\xi_{\alpha,\nu})}
%		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}\\
%	&\hspace{10pt}+ \frac{1}{8\alpha(\alpha - 1)}
%		\left(\frac{\xi_{\alpha,\nu}^{k-2\alpha}e^{-\xi}U(2\alpha+1,1+k-2\alpha,\xi_{\alpha,\nu})}
%		{\Gamma^+(k-2\alpha,\xi_{\alpha,\nu})}-1\right)\\
%	&\hspace{10pt}- \frac{(\alpha - 1/2)^2 \xi_{\alpha,\nu}^2 \Gamma^+(k-2\alpha -2,\xi_{\alpha,\nu})}
%		{4(\alpha - 1)^2 \Gamma^+(k-2\alpha,\xi_{\alpha,\nu})}
%\end{align*}
while for $\alpha = 1$
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

From the exact expression for $c_\infty(k)$ we obtain its asymptotic behavior as $k \to \infty$. For this we first define, for any $\frac{1}{2} < \alpha < \frac{3}{4}$,
\begin{equation}\label{eq:def_C_alpha}
	C_\alpha = \frac{2^{-4\alpha - 1}(3\alpha - 1)}{\alpha(\alpha-1)^2} 
	+ \frac{\alpha - \frac{1}{2}}{2(\alpha - 1)\alpha} B^-(\frac{1}{2},2\alpha + 1, 2\alpha - 2)
	- \frac{1}{4(\alpha - 1)}B(2\alpha, 3\alpha - 4).
\end{equation}
Moreover, to simplify the statement we define the scaling function 
\begin{equation}\label{eq:def_scaling_function}
s_\alpha(k) = \begin{cases} 
		k^{4\alpha-2} &\frac{1}{2}<\alpha<\frac{3}{4}, \\
		\frac{k}{\log(k)} & \alpha = \frac{3}{4}, \\
		k &\alpha > \frac{3}{4}.
\end{cases}
\end{equation}

@@

The following is an immediate consequence of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}.

@@

\PvdH{@All: It would be nice if we could add some simulations showing this scaling in practice, especially the different regimes.}

The majority of the paper is dedicated to prove Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}, which is based on a sequence of subsequent results for the local clustering function in related random graph models and Theorem~\ref{thm:asymptotics_average_clustering_P}. The exact flow of the argument is explained in Section \ref{sec:proof_outline}. The proof for Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}, using these results, can be found in Section \ref{ssec:proof_main_results}. Here we also give the proof of Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The exact result for the local clustering coefficient and function, as well as the proofs of Theorem~\ref{thm:exact_expression_c_infty} and~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}.

@@

\paragraph{Transition in the scaling of local clustering}
%Note that Theorem \ref{thm:asymptotic_local_clustering_hyperbolic} implies that for $\frac{1}{2} < \alpha < \frac{3}{4}$ the local clustering $c_{\H,n}(k)$ decays as $k^{-(4\alpha - 2)}$ while for $\alpha > \frac{3}{4}$ it is $k^{-1}$.

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma [??] \PvdH{Find the reference for this Lemma}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$ we have
\[
	\lim_{n \to \infty} n k_n^{-(2\alpha + 1)} 
	= \lim_{n \to \infty} \left(n^{-\frac{1}{2\alpha + 1}} k_n\right)^{-(2\alpha + 1)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = \lim_{n \to \infty} 1 - \bigT{n k_n^{-(2\alpha + 1)}} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$.

%Maybe, it would make sense to incorporate this statement into theorem \ref{thm:asymptotic_local_clustering_hyperbolic}, i.e. that the asymptotic behaviour of $c(k_n)$ is not well-defined for $k_n \gg n^{\frac{1}{2\alpha+1}}$? Resp. claim that a.a.s. the number of vertices of degree $k_n \gg n^{\frac{1}{2\alpha+1}}$ is zero, such that taking the average over these is not well-defined.

@@

%\paragraph{Local clustering for growing versus fixed degrees}
%%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:asymptotic_local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$. 
%
%\MS{We note that the above statements / theorems are disjoint: in one case the degree is fixed corresponding to a constant sequence and in this case the error only depends on $n$, whereas in the other case, the degrees are assumed to be increasing, so in particular unbounded resp. tending to infinity and only the leading term in the asymptotic expansion w.r.t. $k_n \rightarrow \infty$ is given. This allows to study vertices of relatively high degree, e.g. close to the maximum degree (note that there are a.a.s. no vertices with degree exactly the expected maximum degree of $k_n = n^{\frac{1}{2\alpha}}$).}

%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:asymptotic_local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$.

>>>>

The results are obtained by coupling the hyperbolic random graph to an infinite random graph $G_\Pcal(\alpha,\nu)$, which we define in Section~\ref{ssec:infinite_model}, and show that the limit of the local clustering coefficient and function for the hyperbolic random graph are given by those for the infinite model.

\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function and $B^-(x,a,b)$ the lower incomplete beta-function. We also write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

@@

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to as constant $c_\infty$, which we can explicitly write down in terms of the special function defined above.

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0,
\]
where $c_\infty$ is defined for $\alpha \ne 1$ as
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

@@

For the local clustering function we also obtain convergence in expectation to a function $c_\infty(k)$ which can be explicitly stated using the special functions.

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

From the expression for $c_\infty(k)$ we obtain its asymptotic behavior as $k \to \infty$. For this we first define, for any $\frac{1}{2} < \alpha < \frac{3}{4}$,
\begin{equation}\label{eq:def_C_alpha}
	C_\alpha = \frac{2^{-4\alpha - 1}(3\alpha - 1)}{\alpha(\alpha-1)^2} 
	+ \frac{\alpha - \frac{1}{2}}{2(\alpha - 1)\alpha} B^-(\frac{1}{2},2\alpha + 1, 2\alpha - 2)
	- \frac{1}{4(\alpha - 1)}B(2\alpha, 3\alpha - 4).
\end{equation}
Moreover, to simplify the statement we define the scaling function 
\begin{equation}\label{eq:def_scaling_function}
s_\alpha(k) = \begin{cases} 
		k^{4\alpha-2} &\frac{1}{2}<\alpha<\frac{3}{4}, \\
		\frac{k}{\log(k)} & \alpha = \frac{3}{4}, \\
		k &\alpha > \frac{3}{4}.
\end{cases}
\end{equation}

@@

The following is an immediate consequence of Theorem~\ref{thm:local_clustering_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}.

@@

The majority of the paper is dedicated to prove Theorem~\ref{thm:local_clustering_hyperbolic}, which is based on a sequence of subsequent results for the local clustering function in related random graph models and Theorem~\ref{thm:asymptotics_average_clustering_P}. The exact flow of the argument is explained in Section \ref{sec:proof_outline}. The proof for Theorem~\ref{thm:local_clustering_hyperbolic}, using these results, can be found in Section \ref{ssec:proof_main_results}. Here we also give the proof of Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proof of Theorem~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}.

@@

\paragraph{Transition in the scaling of local clustering}
%Note that Theorem \ref{thm:local_clustering_hyperbolic} implies that for $\frac{1}{2} < \alpha < \frac{3}{4}$ the local clustering $c_{\H,n}(k)$ decays as $k^{-(4\alpha - 2)}$ while for $\alpha > \frac{3}{4}$ it is $k^{-1}$.

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:N_k_HP_P}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n^{-(2\alpha + 1)} n$ tends to zero as $n \to \infty$ we have that 
\[
	\lim_{n \to \infty} \Exp{N_{\H,n}(k_n)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n$ is not $\smallO{n^{\frac{1}{2\alpha + 1}}}$.

%Maybe, it would make sense to incorporate this statement into theorem \ref{thm:local_clustering_hyperbolic}, i.e. that the asymptotic behaviour of $c(k_n)$ is not well-defined for $k_n \gg n^{\frac{1}{2\alpha+1}}$? Resp. claim that a.a.s. the number of vertices of degree $k_n \gg n^{\frac{1}{2\alpha+1}}$ is zero, such that taking the average over these is not well-defined.

@@

%\paragraph{Local clustering for growing versus fixed degrees}
%%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$. 
%
%\MS{We note that the above statements / theorems are disjoint: in one case the degree is fixed corresponding to a constant sequence and in this case the error only depends on $n$, whereas in the other case, the degrees are assumed to be increasing, so in particular unbounded resp. tending to infinity and only the leading term in the asymptotic expansion w.r.t. $k_n \rightarrow \infty$ is given. This allows to study vertices of relatively high degree, e.g. close to the maximum degree (note that there are a.a.s. no vertices with degree exactly the expected maximum degree of $k_n = n^{\frac{1}{2\alpha}}$).}

%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$.
